The Effects of the 2023 Rule Changes in Major League Baseball

### 2023-11-13 manuscript/paper.tex
completed the Results section of manuscript

Insert description about Table~\ref{tab:table1} here.

\begin{table}[tbp]
  \caption{The results from the 9 tests}
  \label{tab:table1}
  \centering
  \begin{tabular}[t]{lccccc}
    \toprule
    & 2023 & 2022 & \multicolumn{2}{c}{P-value}\\
    \cmidrule(lr){4-5}
    &          &                & \(t\)-test & Wilcoxon's\\
    \midrule
    Stolen Bases & 7.226 & 4.994 & 0.0000 & 0.0011 \\ 
    Strikeouts & 85.445 & 81.731 & 0.1596 & 0.0000 \\ 
    Walks & 32.876 & 30.279 & 0.0599 & 0.0791 \\ 
    Batting Avg & 0.243 & 0.237 & 0.0274 & 0.0154 \\ 
    Slugging \% & 0.401 & 0.384 & 0.0008 & 0.0007 \\ 
    BABIP & 0.293 & 0.288 & 0.0793 & 0.0695 \\ 
    xBA & 0.241 & 0.235 & 0.0006 & 0.0005 \\ 
    Hits on GB & 27.345 & 26.399 & 0.3757 & 0.3212 \\ 
    GIDP & 7.232 & 6.925 & 0.3523 & 0.5401 \\ 
    \bottomrule
  \end{tabular}
\end{table}

>>>>

Table~\ref{tab:table1} summarizes the p-values for both the \(t\)-tests and Wilcoxon's
tests for the research questions / hypotheses detailed in Section~\ref{sec:meth}.
The difference in game length test was excluded from this table as this tested
2023 versus 1998-2022 rather than just 2022, due to the data restrictions. Also,
point estimates for each test were provided, one for the 2023 group and one for the
2022 group.

At the top of the table we see that the mean number of stolen bases for each player
in 2023 is 7.226, and 4.994 in 2022 (these are the ``point estimates'') for the tests.
This difference is statistically significant for both the \(t\)-test and Wilcoxon's test
at $\alpha$ = 0.05, the significance level used for all of the tests in the table,
as the p-values are 0.0000 (extremely small, rounded for presentation purposes) and
0.0011 respectively. Adjusting the p-values to account for multiple tests using the
Bonferroni correction also shows significance for both of these tests for stolen bases.
This gives us ample evidence to conclude that the mean number of stolen bases in 2023
significantly increased from 2022.

On the other hand, the number of strikeouts did not seem to significantly change,
as its \(t\)-test p-value is 0.1596. The Wilcoxon's test p-value is extremely small
however, so the two tests are conflicting in their conclusions. Once we look at the
95\% confidence interval (CI) of \((-1.46, 8.89)\) though, we see that it contains zero,
meaning we cannot say the difference in the number of strikeouts year to year is significant.
Moving down the table, other hypotheses that were not significant were Walks, BABIP,
Hits on Ground Balls, and Ground into Double Play (GIDP) for both tests at the
5\% significance level. Hits on Ground Balls was particularly surprising, as
we expected that with no defensive shift anymore, more ground balls would become
base hits.

The difference in Batting Average, Slugging Percentage, and Expected Batting Average (xBA)
from 2022 to 2023 were all statistically significant for both the \(t\)-test and
Wilcoxon's Rank Sum test as their p-values were all less than $\alpha$ = 0.05. However,
using the Bonferroni correction, the adjusted p-values for the Batting Average
tests no longer were significant. So now there is only enough evidence to support
that the difference from 2022 to 2023 in Slugging Percentage and xBA is significant.

Also, for the game time data not shown in Table~\ref{tab:table1}, the average game
length from 1998 to 2022 was 2.92 hours (2 hrs 55 mins), compared to only 2.65 hours
(2 hrs 39 mins) in 2023. The \(t\)-test and Wilcoxon's test for the difference in
these values yielded two extremely small p-values, so we have enough evidence to 
conclude that the average game length is significantly shorter in 2023 compared to
that in the previous 25 years. This was expected, as we assumed the pitch clock would
speed up the game considerably.

\begin{table}[tbp]
  \caption{The results from the 9 baseball statistics tests}
  \label{tab:table1}
  \centering
  \begin{tabular}[t]{lccccc}
    \toprule
    & 2023 & 2022 & \multicolumn{2}{c}{P-value}\\
    \cmidrule(lr){4-5}
    &          &                & \(t\)-test & Wilcoxon's\\
    \midrule
    Stolen Bases & 7.226 & 4.994 & 0.0000 & 0.0011 \\ 
    Strikeouts & 85.445 & 81.731 & 0.1596 & 0.0000 \\ 
    Walks & 32.876 & 30.279 & 0.0599 & 0.0791 \\ 
    Batting Avg & 0.243 & 0.237 & 0.0274 & 0.0154 \\ 
    Slugging \% & 0.401 & 0.384 & 0.0008 & 0.0007 \\ 
    BABIP & 0.293 & 0.288 & 0.0793 & 0.0695 \\ 
    xBA & 0.241 & 0.235 & 0.0006 & 0.0005 \\ 
    Hits on GB & 27.345 & 26.399 & 0.3757 & 0.3212 \\ 
    GIDP & 7.232 & 6.925 & 0.3523 & 0.5401 \\ 
    \bottomrule
  \end{tabular}
\end{table}
